---
title: "Data Import"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


## Data Import: CSVs

let's import this FAS_liters.csv, using 'readr' package

```{r}
litters_df = read_csv("data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df

```

Look at the data

```{r}
litters_df
head(litters_df)
tail(litters_df)
```


```{r, eval=FALSE}
View(litters_df)
#If you have View without eval=FALSE, break the code
summary(litters_df)
skimr::skim(litters_df)
```

`read_csv` options

```{r, eval=FALSE}
read_csv("data/FAS_litters.csv", na = c("","NA", 999, 88), skip = 2)

```

## Read other file

We need to read in a excel spreadsheet
- somtimes, when you open the excel file, will see some $~ excel file, it will create a backup
- you don't see it, but the git sees it
- purpose is so you can revert
- don't commit it. When you close the excel, it will disapear
- act as if it's not there

```{r}
mlb_df = read_excel("data/mlb11.xlsx")

```

```{r, echo = FALSE}
View(mlb_df)
```

```{r}
lotr_words = read_excel("data/lotR_Words.xlsx",
                        range = "B3:D6")
lotr_words
```


## still more formats...

Read in a SAS dataset

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
pulse_df
```


## Data Export

```{r}

write_csv(lotr_words, file = "data/Results/lotr_words.csv")

dont_do_this_df = read.csv("data/FAS_litters.csv")
dont_do_this_df
```

















